---
title: "California Electricity Overview"
format:
  html:
    toc: true
    toc-depth: 3
    toc-collapse: false
    number-sections: false
    fig-align: center
execute:
  freeze: auto
---

# California Electricity Trends

This page gives an overview of California's trends in electricity usage and customers for our three selected sectors, residential, commercial, and industrial. The first plot exhibits the number of customer accounts for each sector and shows the large amount of residential customers in the state. The next plot displays combined heat and power capacity for the commercial and industrial sectors. This plot helps us understand the emphasis of energy efficiency for these two sectors. Together, these plots help paint the picture of California's electricity landscape. 

---

## Background: California Energy Context

California plays a central role in U.S. energy discussions because it is both one of the largest energy consumers and a national leader in renewable energy policy. Its ambitious clean energy goals, including targets for carbon neutrality by 2045 and large-scale adoption of solar and wind power, have reshaped how energy is produced, distributed, and priced across the state. California’s energy sector is characterized by a large residential customer base alongside major industrial and commercial users. These efforts to decarbonize and modernize the grid often contribute to higher electricity prices. A major challenge that California faces is the necessity for massive investment in safer infrastructure to help protect against wildfires. These costs are not only incurred by utilities, but are also passed on to consumers.The data used for this analysis come from the U.S. Energy Information Administration (EIA). 

---

## Customer Accounts by Sector

This visualization shows how customer accounts differ across California’s residential, commercial, and industrial sectors.

```{r}
#| echo: false
#| message: false
#| warning: false

knitr::include_graphics(
here::here("figs", "CA_customer_accounts_plot.png")
)
```

From this combined point and line plot, we can see that there are significantly more residential users than commercial or industrial users, with residential accounts exceeding 13 million compared to fewer than 2 million for the other sectors. The residential trend line remains relatively stable over time, suggesting steady population growth and consistent electricity demand across households. However, the residential data points also display greater variation, indicating fluctuations that may be driven by seasonal demand shifts, changing energy efficiency programs, or adoption of distributed energy resources like rooftop solar. In contrast, commercial and industrial accounts show slower and smaller increases, reflecting the relatively limited number of large-scale energy consumers in California’s economy. The visual gap between the lines emphasizes how dominant the residential sector is in shaping California’s overall energy usage patterns. This composition highlights the importance of residential energy policy in achieving the state’s broader efficiency and sustainability goals.

## California Combined Heat and Power (CHP) for Commercial and Industrial Sector

This visualization shows how combined heat and power differ across California’s commercial and industrial sector.


```{r}
#| echo: false
#| message: false
#| warning: false

knitr::include_graphics(
here::here("figs", "CA_chp_plot.png")
)
```

This plot shows California’s Combined Heat and Power (CHP) net generation by sector from 2004 to 2023. CHP systems simultaneously produce electricity and useful heat from a single fuel source, making them far more energy efficient than separate heat and power generation. The data reveal that CHP generation is overwhelmingly concentrated in the industrial sector, which consistently accounts for the vast majority of total output. This reflects the sector’s heavy energy needs and its economic incentive to adopt efficiency measures that reduce waste heat and energy costs. In contrast, the commercial sector contributes only a small share, indicating that CHP adoption in smaller-scale facilities has been more limited. The gradual decline in total CHP output in recent years may point to shifts toward cleaner, grid-based renewables and changing industrial activity patterns. Overall, the dominance of industrial CHP underscores California’s focus on improving energy efficiency within its largest and most energy-intensive operations.

